---
title: 'Predictive Analytics Exam Module 5 Section 3'

---
If you are using R 3.6.0 or later, the following command will ensure that the random number generator being used is the same as in prior versions of R. Because these modules were developed using R 3.5.0, output that depends on random numbers used the older generator. Running this code will have your output match that shown in the modules.


```{r}
# Run this chunk if using R 3.6.0 or later.
RNGkind(sample.kind = "Rounding")
```

If you are using R 4.0.0 or later, the following command will ensure that read.csv() interprets variables whose values are characters as factor variables. This was the default behavior in prior versions of R. All code is written assuming such variables are factor variables.

```{r}
# Run this chunk if using R 4.0.0 or later. You may get what looks like an error message. It can be ignored.
options(stringsAsFactors = TRUE)
```

Run CHUNK 1 to load the mortality data and ggplot2.

```{r echo = FALSE}
# CHUNK 1
data.mortality <- read.csv("soa_mortality_data.csv")
library(ggplot2)
```

Run CHUNK 2 to check for outliers in the duration variable.

```{r echo = FALSE}
# CHUNK 2
library(gridExtra)
p1 <- ggplot(data = data.mortality, aes(duration, ..density..)) +
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.5)
p2 <- ggplot(data = data.mortality, aes(x = "duration", y = duration)) +
  geom_boxplot(fill = "blue", alpha = 0.5)
grid.arrange(p1, p2, ncol = 2)
```

In CHUNK 3 modify the code from CHUNK 2 to check for outliers in the duration variable after a log transformation.

```{r echo = FALSE}
# CHUNK 3
p1 <- ggplot(data = data.mortality, aes(duration, ..density..)) +
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.5)
p2 <- ggplot(data = data.mortality, aes(x = "duration", y = duration)) +
  geom_boxplot(fill = "blue", alpha = 0.5)
grid.arrange(p1, p2, ncol = 2)
```

CHUNK 4 has the modified code.

```{r echo = FALSE}
# CHUNK 4
data.mortality$durationLog <- log(data.mortality$duration)
p1 <- ggplot(data = data.mortality, aes(durationLog, ..density..)) +
  geom_histogram(binwidth = 0.1, fill = "blue", alpha = 0.5)
p2 <- ggplot(data = data.mortality, aes(x = "duration", y = durationLog)) +
  geom_boxplot(fill = "blue", alpha = 0.5)
grid.arrange(p1, p2, ncol = 2)
```

Run CHUNK 5 to see the effect of an outlier on a regression line.

```{r echo = FALSE}
# CHUNK 5
# The x-values are uniform(0,10) random numbers
# The y-values are the x-values plus a standard normal variable.
sample.x <- runif(10, min = 0, max = 10)
sample.y <- sample.x + rnorm(10, 0, 1)

# The true regression line is y=x. The fitted line to the random data should be close to this.
line.real <- lm(sample.y ~ sample.x)

# Now change the tenth y-value to 50, an outlier.
sample.y[10] <- 50

# Redo the regression.
line.bias <- lm(sample.y ~ sample.x)

# Plot the two regression lines.
plot(sample.y ~ sample.x)
abline(line.real, col = "blue")
abline(line.bias, col = "red")
```

Run CHUNK 6 to revisit the duration variable on a log scale.

```{r echo = FALSE}
# CHUNK 6

p1 <- ggplot(data = data.mortality, aes(x = "duration", y = duration)) +
  geom_boxplot(fill = "orange", alpha = 0.5) +
  scale_y_log10()
p2 <- ggplot(data = data.mortality, aes(x = prodcat, y = duration)) +
  geom_boxplot(fill = "orange", alpha = 0.5) +
  scale_y_log10()
grid.arrange(p1, p2, ncol = 2)
```

CHUNK 7 makes a graph that illustrates the difference between squared and absolute error. It is provided here in case you are interested in how the graph was made.

```{r echo = FALSE}
# CHUNK 7

x <- seq(-1.5, 1.5, 0.1)
errors <- data.frame(
  error = x,
  squared = x^2,
  abs = abs(x)
)

ggplot(data = errors, aes(x = error)) +
  geom_line(aes(y = squared), color = "red") +
  geom_line(aes(y = abs), color = "blue") +
  geom_line(aes(y = 1)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 1, fill = "yellow", alpha = 0.2) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = 1, ymax = Inf, fill = "orange", alpha = 0.2)
```

Run CHUNK 8 to examine duration versus actual face for term insurance products.

```{r}
# CHUNK 8
p1 <- ggplot(data = subset(data.mortality, prodcat == "TRM"), aes(x = duration, y = actual_face)) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(limits = c(0, 80)) +
  scale_y_log10(limits = c(1, 1.0e+07)) +
  annotate("rect", xmin = 39, xmax = 43, ymin = 600, ymax = 10000, fill = "red", alpha = 0.2)
p1
```

Run CHUNK 9 to see an example where one variable is categorical and the other is numeric.

```{r}
# CHUNK 9
p1 <- ggplot(data = subset(data.mortality, actual_cnt >= 1), aes(x = smoker, y = issage, fill = smoker)) +
  geom_boxplot()
p2 <- ggplot(data = subset(data.mortality, actual_cnt >= 1), aes(x = "smoker", y = issage)) +
  geom_boxplot()
grid.arrange(p1, p2, ncol = 2)
```

Run CHUNK 10 to make a frequency table of distribution channel and smoking status.

```{r}
# CHUNK 10
tbl <- table(data.mortality$distchan, data.mortality$smoker)
tbl
```
